---
title: "Lab 02"
author: "Danah Milyani"
subtitle: "CS3172-1, Spring 2023, Effat University"
output:
  word_document: default
  pdf_document: default
---

# Packages

```{r load-packages, message = FALSE, warning = FALSE}
library(tidyverse) 
library(scales)
``` 

# Data

```{r read-in-data, message = FALSE}
cas <- read_rds("data/canada_survey.rds")
```

# Tasks

## Task 1

```{r}
cas <- filter(cas,energy_expense > 0,household_income>0)
```

```{r}
cas <- mutate(cas,marital_status = factor(marital_status))
```

## Task 2

```{r}
cas <- mutate(cas,heat_equip = case_when( heat_equip == 1 ~ 'steam' , 
                                           heat_equip == 2 ~ 'forced air' ,
                                           heat_equip == 3 ~ 'stove' ,
                                           heat_equip == 4 ~ ' electric heating' ))
```

```{r}
cas <- mutate(cas,heat_fuel = case_when( heat_fuel == 1 ~ 'oil' , 
                                           heat_fuel == 2 ~ 'gas' ,
                                          heat_fuel == 3 ~ 'electricity' ,
                                           heat_fuel == 4 ~ ' other' ))
```

## Task 3

```{r}
options(dplyr.summarise.inform = FALSE)
cas1 <- cas %>% group_by(heat_equip,heat_fuel)  %>% summarise( mean_ener_exp = mean(energy_expense),
                                                              median_ener_exp = median(energy_expense),
                                                              sd_ener_exp = sd(energy_expense))
```

- Provide the answer to the theoretical questions here

o	What combination of fuel type and equipment has the highest average energy expense?
the combination of fuel type and equipment that has the highest average energy expense are, stove and oil 
o	Which combination has the most variability with regards to energy expense?
The most variability is oil and forced air
o	Which type of heating equipment doesn’t take all possible fuel types?
The type heating equipment that doesn’t take all possible fuel types is electric heating

## Task 4

```{r}
ggplot(cas,mapping=aes(x=heat_fuel, y=energy_expense)) + geom_boxplot() + 
  scale_y_continuous(labels = scales::dollar_format()) +
  facet_wrap(~ heat_equip, nrow=2) + labs(title =' energy expense vs heating type' ,
    x= "heat fueling type" , y = "energy expense" )
```

## Task 5

```{r}


cas2 <- cas %>% mutate(energy_prop = energy_expense/household_income) %>%
  arrange(desc(energy_prop)) %>% slice(1,n()) %>% glimpse()
```

## Task 6

```{r message=FALSE}

cas %>% group_by(year,province) %>%  summarise( median_energy_per_room =
    median(energy_expense/rooms)) %>% slice_min(median_energy_per_room)
```

## Task 7

```{r fig.width=7, fig.height=6, message=FALSE}

cas %>% mutate(energy_prop = energy_expense / household_income,
  vulnerable = if_else(energy_prop > 0.05, "vulnerable", "not")) %>%
  group_by(education, age) %>%
  summarize(prop_vulnerable = mean(vulnerable == "vulnerable")) %>%
  ungroup() %>%
  ggplot(aes(x = education, y = age, fill = prop_vulnerable)) + 
  geom_raster() + 
  scale_fill_gradient(low = "white", high = "red") +
  labs(x = "Education", y = "Age", fill = "Proportion\n energy \n vulnerable") +
  theme_bw()

```

## Task 8


